---
title: "Prediction Assignmen Writeup"
author: "Reginald W Carroll"
date: "May 24, 2018"
output: md_document
---


## Practical Machine Learning - Pediction Assignment Writeup
Background:

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

## Install packages for the project

```{r}
library(tidyverse)
library(caret)
library(randomForest)
```


## Load Data (pml-training.csv and pml-testing.csv) 
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har


```{r}
url.trainData <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url.testData  <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training      <- read.csv(url(url.trainData), na.strings = c("NA", "", "#DIV0!"))
testing       <- read.csv(url(url.testData), na.strings = c("NA", "", "#DIV0!"))
set.seed(23)
```
## Cleansing the data
```{r}
train.noNA <- training[colSums(!is.na(training))>(nrow(training)*.9)]
test       <- testing[colSums(!is.na(testing))>(nrow(testing)*.9)]
```
## Remove the first 7 columns because they are not needed for calculation.
```{r}
train.cleansese <- train.noNA[,-c(1:7)]
```

## Removing highly correlated predictors. Get only numeric columns
 
```{r}
num.cols <-sapply(train.cleansese,is.numeric)
```
## Create a correlation matrix

```{r}
cor.data <- cor(train.cleansese[,num.cols])
```
## Identify correlated predictors for removal
```{r}
high.cor <- findCorrelation(cor.data, cutoff = .75)
train.cor <- train.cleansese[,-high.cor]
train.cor$classe <- as.factor(train.cor$classe)
```

## Split the data into training (75%) and testing (25%) datasets to be used for Cross Validation 
```{r}
inTrain <- createDataPartition(y=train.cor$classe,p=0.75,list=FALSE)
training <- train.cor[inTrain,]
testing <- train.cor[-inTrain,]
```
## Plot the training dataset
```{r}
qplot(classe, data=training,  main = "Activity Classes (Training Data)")
```

## Train the model using RandomForest
```{r}
modelfit.rf <- randomForest(classe ~.,training)
pred.rf <- predict(modelfit.rf,testing)

```
## View the Confusion Matrix
Based on the Confusion Matrix summary, Sensitivity is .9993 and Specificity is .9991 the balance accuracy is .9992

```{r}
cmatrix <- confusionMatrix(pred.rf,testing$classe)
print(cmatrix)
```

## Plot model fit
```{r}
plot(modelfit.rf)
```
## Final prediction using the test dataset (pml-testing.csv)
```{r}
final.pred <- predict(modelfit.rf, test, type = "class")
print(final.pred)
```
## Plot the Final predictiion
```{r}
qplot(final.pred, main = "Activity Classes (Test Data)")
```
